---
title: "Customising and extending your cost model"
author: "Robin Linacre"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Example of simple cost model}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r results='hide', message=FALSE, warning=FALSE}
library(DT)

```

## Introduction

This vignette illustrates how to customise and extend your cost model.

## Adding settings

### Categorisation columns

By default, each table of cost assumptions has three columns, `category_1,category_2 and category_3`.  This enables the user to breakdown final results at varying degrees of granularity.

For example, oneoff cost assumptions may look like this:

```{r, echo=FALSE}

df <- readr::read_csv(system.file("extdata", "oneoff_costs_1.csv", package="costmodelr"), col_types=readr::cols())
DT::datatable(df, options=list(scrollX = TRUE))
```

However, the user can provide additional categorisation columns, or overwrite these defaults.  For example, if in addition to `category_1,category_2 and category_3`, the user wants to provide a forth column called `additional breakdown`, they should write the following:

```{r, eval=FALSE}
cost_moddel <- create_cost_model(key_dates) %>%
                  setting_append_to_categorisation_columns("additional_breakdown")
```

If the user wanted instead to provide two categorisation columns, `cat_1` and `cat_2` (and not to provide `category_1,category_2 and category_3`), they might instead write:

```{r, eval=FALSE}
cost_moddel <- create_cost_model(key_dates) %>%
                  setting_categorisation_columns(c("cat_1", "cat_2"))
```

### Deflator base date

The user can specify the 'base date' for real costs.  For instance if the user wants all costs to be quoted in '2017 pounds', they would specify the following:

  
    
```{r, eval=FALSE}
cost_model <- create_cost_model(key_dates) %>%
    setting_deflator_base_date(as.Date("2017-01-01"))
```
    
## Full example

The following provides a full example.  The two assumptions files are reproduced below.

```{r, eval=FALSE}
key_dates <- readr::read_csv(system.file("extdata", "key_dates_2.csv", package="costmodelr")))
oneoff <- readr::read_csv(system.file("extdata", "oneoff_costs_2.csv", package="costmodelr")))
recurring <- readr::read_csv(system.file("extdata", "recurring_cost_2.csv", package="costmodelr")))

cost_model <- create_cost_model(key_dates) %>%
  setting_append_to_categorisation_columns("additional_breakdown") %>%
  setting_deflator_base_date(as.Date("2017-01-01")) %>%
  add_oneoff_costs(oneoff) %>%
  add_recurring_cost(recurring) %>%
  run_cost_model()
```

`oneoff_costs_2.csv`

```{r, echo=FALSE}
df <- readr::read_csv(system.file("extdata", "oneoff_costs_2.csv", package="costmodelr"), col_types=readr::cols())
DT::datatable(df, options=list(scrollX = TRUE))
```

`recurring_cost_2.csv`

```{r, echo=FALSE}
df <- readr::read_csv(system.file("extdata", "recurring_cost_2.csv", package="costmodelr"), col_types=readr::cols())
DT::datatable(df, options=list(scrollX = TRUE))
```
